Post-replication citation patterns in psychology: Four case studies

### 2021-07-14 manuscript/appendix.Rmd
updated after round 2 revisions

```{r protocolAmendments}
tibble(
  "Issue" = c("Sampling frame", "Definition of pre-replication assessment period", "Exclusion of replication year from qualitative assessment", "Automated extraction of citation contexts and replication co-citation", "Counter-arguments", "Dual-coding for citation valence classification", "Examination of characteristics of articles providing counter-arguments.","Renaming of two citation patterns", "Data exclusions"),
  "Amendment" = c("The end date of the sampling frame was initially April 6th 2018 (see pre-registered protocol). Later, we decided to extend the sampling frame to April 26th 2019 to learn more about how the citation curves developed over time (see amended protocol). Finally, we decided to extend the sampling frame to December 31st 2019 to ensure we had all citations for the year 2019.", "Initially, the pre-replication assessment period was defined as 2015 (i.e., replication year - 1) for the Registered Replication Report cases and all pre-replication years for the other cases. In practice, when the replication year - 1 definition is used for all cases, only one study (Caruso) has a citing article outside of this period, and it is only a single article. Therefore, we have decided that it is more straightforward to use the same definition of the pre-replication assessment period (replication year - 1) for all cases.", "The pre-registered protocol neglected to mention that we intended to exclude the replication year from qualitative assessment. This change was documented in the amended protocol.", "Initially, we intended to use Matlab code to partly automate the extraction of ‘citation contexts’ (text surrounding in-text citations) and identify replication co-citation. However, we found that the size of the relevant citation context varied substantially between articles, so we decided to extract all citation contexts manually. We also decided to identify replication co-citation manually too. These changes were documented in the amended protocol.","In the amended protocol, we clarified that we would attempt to identify counter-arguments in all articles that cited the original study and the replication study.", "Initially, we planned to dual code small batches of articles depending on the degree of inter-rater reliability. However, after encountering a number of somewhat ambiguous cases in our initial coding, we decided to dual-code all articles in order to reduce subjectivity and increase consistency between coders. This is documented in the amended protocol.", "After completing the study we thought it would be informative to examine the article types and journals of articles that provided counter-arguments, and any overlap in authorship with the original study authors or prior collaborators of the first authors of the original studies. These analyses were not pre-registered.", "After completing the study we decided to revise the following terminology from the protocol in order to improve clarity: “principled defence” was changed to “explicit defence” and “hollow defence” was changed to “absent defence”.", "The criteria to exclude data from qualitative assessment (no access, non-English language, and no original study citation in the main text, were not pre-registered. These exclusions were necessary because they precluded the qualitative assessment.")
) %>%
  kable(booktabs=T,longtable=T,caption = "Protocol amendments.") %>%
  column_spec(1,width='4cm') %>%
  column_spec(2,width='11cm')
```

@@

>>>>

```{r protocolAmendments}
tibble(
  "Issue" = c("Sampling frame", "Definition of pre-replication assessment period", "Exclusion of replication year from qualitative assessment", "Automated extraction of citation contexts and replication co-citation", "Counter-arguments", "Dual-coding for citation valence classification", "Examination of characteristics of articles providing counter-arguments.","Renaming of two citation patterns", "Data exclusions", "Citations to replication studies and co-citation of original studies", "Additional access attempts and translation"),
  "Amendment" = c("The end date of the sampling frame was initially April 6th 2018 (see pre-registered protocol). Later, we decided to extend the sampling frame to April 26th 2019 to learn more about how the citation curves developed over time (see amended protocol). Finally, we decided to extend the sampling frame to December 31st 2019 to ensure we had all citations for the year 2019.", "Initially, the pre-replication assessment period was defined as 2015 (i.e., replication year - 1) for the Registered Replication Report cases and all pre-replication years for the other cases. In practice, when the replication year - 1 definition is used for all cases, only one study (Caruso) has a citing article outside of this period, and it is only a single article. Therefore, we have decided that it is more straightforward to use the same definition of the pre-replication assessment period (replication year - 1) for all cases.", "The pre-registered protocol neglected to mention that we intended to exclude the replication year from qualitative assessment. This change was documented in the amended protocol.", "Initially, we intended to use Matlab code to partly automate the extraction of ‘citation contexts’ (text surrounding in-text citations) and identify replication co-citation. However, we found that the size of the relevant citation context varied substantially between articles, so we decided to extract all citation contexts manually. We also decided to identify replication co-citation manually too. These changes were documented in the amended protocol.","In the amended protocol, we clarified that we would attempt to identify counter-arguments in all articles that cited the original study and the replication study.", "Initially, we planned to dual code small batches of articles depending on the degree of inter-rater reliability. However, we encountered a number of somewhat ambiguous cases in our initial coding which suggested there may be more subjectivity involved in the classification of citation valence than we had originally anticipated. For example, an article may contain several favourable statements and several unfavourable statements, in which case one needs to determine how to appropriately weight those statements in order to arrive at a classification that accurately represents the author’s viewpoint. To address this, we decided to minimize subjectivity by dual-coding all articles, with disagreements resolved through discussion (with a third coder if necessary). This is documented in the amended protocol.", "After completing the study we thought it would be informative to examine the article types and journals of articles that provided counter-arguments, and any overlap in authorship with the original study authors or prior collaborators of the first authors of the original studies. These analyses were not pre-registered.", "After completing the study we decided to revise the following terminology from the protocol in order to improve clarity: “principled defence” was changed to “explicit defence” and “hollow defence” was changed to “absent defence”.", "The criteria to exclude data from qualitative assessment (no access and no original study citation in the main text, were not pre-registered. These exclusions were necessary because they precluded the qualitative assessment.","Upon reviewer request, we obtained citation counts for replication studies and checked if they co-cited the relevant original study. These results are presented in the main results section.", "Upon reviewer request, we took additional steps to obtain articles that we were previously unable to access, and to translate articles that were written in a non-English language. The newly accessible and translated articles then underwent qualitative analyses. These new data had minimal impact on the results, we have therefore incorporated them into the main paper. Supplementary Information D describes the process in detail and also contains copies of the original graphs before the data from the newly accessible and translated articles were incorporated.")
) %>%
  kable(booktabs=T,longtable=T,caption = "Protocol amendments.") %>%
  column_spec(1,width='4cm') %>%
  column_spec(2,width='11cm')
```

@@

# Additional access attempts and translation {#appD}

```{r}
d_round2access <- read_csv(here('data','processed','d_round2access.csv'))
```

Initially, we excluded 64 of 632 articles from the qualitative analysis because (1) we could not access the full text via at least two of our institutional library affiliations (n = 58); or (2) they were non-English language (n = 6). Following requests from a reviewer, we took additional steps to access and translate articles (as described in the methods section).

Our additional access attempts were successful for `r d_round2access %>% filter(access == T) %>% nrow()` articles. `r d_round2access %>% filter(access == T) %>% count(source) %>% filter(source == "UVA search") %>% pull(n)` were obtained via Institutional Library search, `r d_round2access %>% filter(access == T) %>% count(source) %>% filter(source == "general search") %>% pull(n)` were obtained via general Internet search, `r d_round2access %>% filter(access == T) %>% count(source) %>% filter(source == "email") %>% pull(n)` were obtained via author e-mail requests, and `r d_round2access %>% filter(access == T) %>% count(source) %>% filter(source == "email; inter-library loan") %>% pull(n)` was obtained via Interlibrary Loan from the University of Illinois. The `R d_round2access %>% filter(access == F) %>% nrow()` remaining articles could not be accessed despite trying all of these methods and contacting a librarian for assistance.

`r d_round2access %>% filter(translationNeeded == T) %>% nrow()` articles were written in non-English language. We translated the relevant citation contexts to English. For `r d_round2access %>% filter(translationNote == "translated by SC from German (SC is native speaker)") %>% nrow()` articles written in German, translation was performed by one of the authors (SC) who is a native German speaker. For the other `r d_round2access %>% filter(translationNote != "translated by SC from German (SC is native speaker)") %>% nrow()` articles, written in Czech, Spanish, Korean, and Russian, we used Google Translate.

The newly accessible and translated articles underwent qualitative analysis according to same procedures as other articles. The new data was incorporated into the main analysis, the results of which are presented in the main body of the paper. The new data had minimal impact on the results.

### 2021-07-14 manuscript/manuscript.Rmd
updated after round 2 revisions

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes

@@

output            : 
  papaja::apa6_pdf:
    includes:
      after_body: "appendix.tex"
---

@@

Table \@ref(tab:patterns) outlines several citation patterns that might follow a contradictory replication result, each reflecting different types of response by the research community. We have tentatively categorised these response patterns as being ‘progressive’ or ‘regressive’ depending on their expected impact on the accumulation of scientific knowledge[^footnote2]. The first set of patterns, belief correction/perpetuation, refer to what is often considered a primary functional role of (contradictory) replication studies - to change belief in the credibility of exaggerated or erroneous original findings (Ioannidis, 2012; Vazire & Holcombe, 2020; Zwaan et al., 2018). In the absence of an explicit defence of an original study that convincingly explains a strongly contradictory replication result (see ‘explicit defence’ below), a progressive response might involve a decrease in favourable citations and an increase in unfavourable citations, reflecting updated beliefs about the credibility of the original finding. Conversely, a regressive response might involve maintenance of (or even increase in) favourable citations, and relatively few unfavourable citations, suggesting a perpetuation of belief in the credibility of the original finding despite the contradictory replication result. Prior research has documented how favourable citations to observational epidemiology studies can persist despite the claims of those studies being strongly contradicted in subsequent randomized trials (Tatsioni et al., 2007). Similarly, it has been reported that even when articles are retracted, they can continue to receive favourable citations (Budd et al., 1998; Fernandez & Vadillo, 2020). Thus, there is evidence that belief in the credibility of original findings can perpetuate even when subsequent events cast doubt on their credibility; however, we are unaware of similar evidence in the context of replication studies that were explicitly designed to test the reliability of prior findings.

@@

The study protocol (rationale, methods, and analysis plan) was pre-registered on April 7th 2018 (https://osf.io/eh5qd/). An amended protocol was registered part way through data collection on May 1st 2019, primarily because we extended the sampling frame to cover additional months (https://osf.io/pdvb5/). All deviations from these protocols are explicitly acknowledged in Appendix \@ref(appA). All data exclusions and measures conducted during this study are reported.

@@

For each citing article undergoing qualitative assessment, we attempted to retrieve the full text via at least two of the institutional libraries we are affiliated with. Inaccessible articles were excluded. For articles for which we could obtain the full text, we classified the research design according to the categories in Table \@ref(tab:articleTypes) and recorded whether the replication study was cited after manual inspection of the reference section (see Table \@ref(tab:patterns): citation balance/bias).

@@

@@

```{r}
excluded_n <- d_contentAnalysis %>%
  filter(excluded == T) %>%
  nrow()
excluded_reason <- d_contentAnalysis %>%
  filter(excluded == T) %>%
  count(exclusionReason)
noAccess_n <- excluded_reason %>%
  filter(exclusionReason == "No access") %>%
  pull(n)
nonEnglish_n <- excluded_reason %>%
  filter(exclusionReason == "Non-English language") %>%
  pull(n)
refCiteOnly_n <- excluded_reason %>%
  filter(exclusionReason == "Cited in references but not in text") %>%
  pull(n)
noCite_n <- excluded_reason %>%
  filter(exclusionReason == "Does not cite target or replication") %>%
  pull(n)
notexcluded_n <- d_contentAnalysis %>%
  filter(excluded == F) %>%
  nrow()
```

In total, `r d_citations %>% nrow()` articles cited one of the original studies of which `r d_contentAnalysis %>% nrow()` articles (after taking a 40% random sample in the Baumeister case) fell within the time period designated for qualitative assessment. Of these `r d_contentAnalysis %>% nrow()` articles, we excluded `r excluded_n` from the qualitative analysis because (1) we could not access the full text (n = `r noAccess_n`); (2) they were non-English language (n = `r nonEnglish_n`); (3) they included a citation to the original study in the reference section, but not in the main text (n = `r refCiteOnly_n`); or (4) manual inspection indicated that they did not actually appear to cite the original study at all (n = `r noCite_n`). Article type classifications for the remaining `r notexcluded_n` articles included in the qualitative analysis are shown in Table \@ref(tab:articleTypes).

@@

Discussion
It has been proposed that replication studies can facilitate scientific self-correction by modifying scientists’ belief in the credibility of published findings (Ioannidis, 2012; Vazire & Holcombe, 2020; Zwaan et al., 2018); however, the extent to which this occurs in practice is unclear. In this study, we investigated how the research community responded to four strongly contradictory replications in the field of psychology by examining post-replication citation patterns for original studies. We observed some ‘progressive’ response patterns in the form of modest active correction patterns (a small decline in favourable citations and a small increase in unfavourable citations), but also even more prominent ‘regressive’ response patternsß in the form of unchallenged belief perpetuation (sustained levels of favourable citations and few unfavourable citations, particularly in the Baumeister case) and considerable citation bias (neglecting to cite the replication study; in all cases aside from Sripada). When authors cited the original study favourably despite the replication result, only half of the articles provided any explicit counter arguments in defence of the original study (an explicit defence). Overall, these findings are consistent with prior observations that favourable citation patterns appear relatively unperturbed by subsequent publication of contradictory results in studies with lower risk of bias (Tatsioni et al., 2007) or even by full retraction (Budd et al., 1998; Fernandez & Vadillo, 2020), and that positive (supportive) evidence is preferentially cited relative to negative (non-supportive) evidence once a theory gets entrenched, despite overwhelming evidence against it (Bastiaansen et al., 2015; Greenberg, 2009).

It is reasonable to question whether the replication results in these case studies should (rationally) have instigated belief change in the research community and thus triggered a more sizable decline in favourable citations than we observed. It is important to note that the case studies we selected were deliberately chosen because the replication results were superior to the original results in terms of both credibility and evidential value. The replications were pre-registered, muli-laboratory studies with large sample sizes. By contrast, the original studies had much smaller sample sizes, and arguably had a much higher risk of bias, as they were not pre-registered, were performed by single teams, and arose in domains affected by publication bias and other questionable research practices (E. C. Carter et al., 2015; Coles et al., 2019; Vadillo, 2019; Vadillo et al., 2016). Thus, it seems reasonable to suggest that the compelling replication results should have reduced belief in the credibility of the original results. However, one could also contend that if there were some flaw in the replication study that undermined its validity (Brandt et al., 2014; Vazire et al., 2021; Fabrigar et al., 2020), then this may provide justification to continue favourably citing the original study. Indeed, in these particular case studies, the validity of the replication results has been challenged by proponents of the original findings (Table 5; Baumeister & Vohs, 2016; Strack, 2016). As this debate remains far from settled (Coles et al., 2019; Vadillo, 2019), ideally any favourable citation of the original studies should at a minimum be accompanied by co-citation of the replication results and some discussion of the discrepant findings.

@@

We have only been able to gauge reactions to replications to the extent that they are reflected in relatively short-term citation patterns. A potential explanation of the apparently cursory treatment of the replication studies could be that researchers only became aware of them after their own research projects had begun and or even had been completed. If one of the original studies had been a key motivator for one’s own study, then it may be difficult to accommodate the strongly contradictory replication results. It may even be tempting to ignore them, or give them only superficial treatment. Moreover, researchers who are convinced that the replication study has squarely refuted the original, may no longer be interested in doing research on a topic where they see no future potential. Additionally, examination of citation patterns would not detect if there had been a correction effect amongst individuals who would not typically cite the original study, such as students, members of the public, or researchers working in other fields. A contested study may continue to be cited favourably by its proponents who remain working in the field. This will suffice to create belief perpetuation in the published literature, even though other scientists may simply no longer be interested in getting involved with such a strongly contested research topic. Relatedly, we did not examine citation patterns beyond 3-5 years post-replication. Some perspectives envision scientific self-correction unfolding over a much longer time scale (Lauden, 1981; Peterson & Panofsky, 2020) and suggest the process is characterised less by the impact of individual study results and more by the gradual accumulation of converging evidence, gradual revision of theoretical understanding, and/or informal sociological processes (e.g., researchers choosing alternative topics to study). Thus, whilst the current findings may contradict the expectations of a more direct and expedited view of the corrective impact of replication studies (Ioannidis, 2012; Vazire & Holcombe, 2020), they are not necessarily inconsistent with a slower and more indirect process of self-correction. Future research could employ longitudinal designs or older historical case studies in order to evaluate citation patterns unfolding over a longer time scale.

@@

```{r render_appendix, include=FALSE}
render_appendix("appendix.Rmd")
```

>>>>

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
always_allow_html: true

@@

output            : 
  papaja::apa6_word
---

@@

Table \@ref(tab:patterns) outlines several citation patterns that might follow a contradictory replication result, each reflecting different types of response by the research community. We have tentatively categorised these response patterns as being ‘progressive’ or ‘regressive’ depending on their expected impact on the accumulation of scientific knowledge[^footnote2]. The first set of patterns, belief correction/perpetuation, refer to what is often considered a primary functional role of (contradictory) replication studies - to change belief in the credibility of exaggerated or erroneous original findings (Ioannidis, 2012; Vazire & Holcombe, 2020; Zwaan et al., 2018). In the absence of an explicit defence of an original study that convincingly explains a strongly contradictory replication result (see ‘explicit defence’ below), a progressive response might involve a decrease in favourable citations and an increase in unfavourable citations, reflecting updated beliefs about the credibility of the original finding. Conversely, a regressive response might involve maintenance of (or even increase in) favourable citations, and relatively few unfavourable citations, suggesting a perpetuation of belief in the credibility of the original finding despite the contradictory replication result. Prior research has documented how favourable citations to observational epidemiology studies can persist despite the claims of those studies being strongly contradicted in subsequent randomized trials (Tatsioni et al., 2007). Similarly, it has been reported that even when articles are retracted, they can continue to receive favourable citations (Budd et al., 1998; Fernandez & Vadillo, 2020). Thus, there is evidence that belief in the credibility of original findings can perpetuate even when subsequent events cast doubt on their credibility; however, we are unaware of similar evidence in the context of studies that were explicitly designed to test the replicabiity of prior findings.

@@

The study protocol (rationale, methods, and analysis plan) was pre-registered on April 7th 2018 (https://osf.io/eh5qd/). An amended protocol was registered part way through data collection on May 1st 2019, primarily because we extended the sampling frame to cover additional months (https://osf.io/pdvb5/). All deviations from these protocols are explicitly acknowledged in Appendix \@ref(appA). We report how we determined our sample size, all data exclusions, all manipulations, and all measures in the study.

@@

For each citing article undergoing qualitative assessment, we attempted to retrieve the full text via several methods in the following order: (1) search of at least two of the institutional libraries we are affiliated with; (2) general Internet search for the article title, including the Google and Google Scholar search engines and Research Gate; (3) e-mail requests to the corresponding author; (4) inter-library loan request. Articles that remained inaccessible after all of these methods were exhausted were excluded. Articles written in non-English language were translated by one of the authors or using Google Translate (see Supplementary Information D). For articles for which we could obtain the full text, we classified the research design according to the categories in Table \@ref(tab:articleTypes) and recorded whether the replication study was cited after manual inspection of the reference section (see Table \@ref(tab:patterns): citation balance/bias).

@@

### Citations to replication studies and co-citation of original studies

Upon reviewer request, we examined citation counts for replication studies and checked whether citing articles also co-cited the relevant original study. To obtain the data, we downloaded bibliographic records from the Clarivate Analytics Web of Science Core Collection accessed via the University of Amsterdam on 16th April, 2021 for articles that cited each replication study (up to the study endpoint - December, 2019) and cross-checked them with our sample of articles that cited the original study. As shown, the replication studies also have a life of their own and they are often cited independently of the specific original study. Often this is easy to explain, e.g., Klein et al. replicated 13 original studies, not just the two that were of interest in our analysis. These studies are also likely to have accrued citations by virtue of being amongst the first highly prominent examples of preregistered multi-laboratory replications in psychology.

```{r}
# identify how many studies citing the replication also cite the original study

## extract citing articles for replication studies and original studies
d_kleinCites <- d_repCitations %>% filter(case == 'klein')
d_wagenmakersCites <- d_repCitations %>% filter(case == 'wagenmakers')
d_haggerCites <- d_repCitations %>% filter(case == 'hagger')
d_baumeisterCites <- d_citations %>% filter(case == 'baumeister')
d_sripadaCites <- d_citations %>% filter(case == 'sripada')
d_carusoCites <- d_citations %>% filter(case == 'caruso')
d_carterCites <- d_citations %>% filter(case == 'carter')
d_strackCites <- d_citations %>% filter(case == 'strack')

## compute citation counts for replication studies
hagCites <- nrow(d_haggerCites)
wagenCites <- nrow(d_wagenmakersCites)
kleinCites <- nrow(d_kleinCites)

## compute how many articles citing the replication study also cite the relevant original study
hagBauCocites <- sum(d_haggerCites$UT %in% d_baumeisterCites$UT)
hagSriCocites <- sum(d_haggerCites$UT %in% d_sripadaCites$UT)
wagenStrackCocites <- sum(d_wagenmakersCites$UT %in% d_strackCites$UT)
kleinCartCocites <- sum(d_kleinCites$UT %in% d_carterCites$UT)
kleinCaruCocites <- sum(d_kleinCites$UT %in% d_carusoCites$UT)

## build table
repCociteTable <- data.frame(
  "Replication" = c("Hagger et al. (2016)", "Klein et al. (2014)", "Wagenmakers et al. (2016)"),
  "Citation" = c(hagCites, kleinCites, wagenCites),
  "Cocitations" = c(paste0(hagBauCocites," (Baumeister et al. 1998) ",hagSriCocites," (Sripada et al. 2014)"), paste0(kleinCartCocites," (Carter et al. 2011) ",kleinCaruCocites," (Caruso et al. 2012)"), paste0(wagenStrackCocites," (Strack et al. 1988)"))
  )
```

```{r}
repCociteTable %>% kable(col.names = c("Replication study", "Citation count","Cocitations of original study"), caption = 'Citations counts for replication studies and co-citation counts to relevant original studies.', booktabs=T)
```

@@

```{r}
excluded_n <- d_contentAnalysis %>%
  filter(excluded == T) %>%
  nrow()
excluded_reason <- d_contentAnalysis %>%
  filter(excluded == T) %>%
  count(exclusionReason)
noAccess_n <- excluded_reason %>%
  filter(exclusionReason == "noAccess") %>%
  pull(n)
refCiteOnly_n <- excluded_reason %>%
  filter(exclusionReason == "Cited in references but not in text") %>%
  pull(n)
noCite_n <- excluded_reason %>%
  filter(exclusionReason == "Does not cite target or replication") %>%
  pull(n)
notexcluded_n <- d_contentAnalysis %>%
  filter(excluded == F) %>%
  nrow()
```

In total, `r d_citations %>% nrow()` articles cited one of the original studies of which `r d_contentAnalysis %>% nrow()` articles (after taking a 40% random sample in the Baumeister case) fell within the time period designated for qualitative assessment. Of these `r d_contentAnalysis %>% nrow()` articles, we excluded `r excluded_n` from the qualitative analysis because (1) we could not access the full text (n = `r noAccess_n`); (2) they included a citation to the original study in the reference section, but not in the main text (n = `r refCiteOnly_n`); or (3) manual inspection indicated that they did not actually appear to cite the original study at all (n = `r noCite_n`). Article type classifications for the remaining `r notexcluded_n` articles included in the qualitative analysis are shown in Table \@ref(tab:articleTypes).

@@

Discussion
It has been proposed that replication studies can facilitate scientific self-correction by modifying scientists’ belief in the credibility of published findings (Ioannidis, 2012; Vazire & Holcombe, 2020; Zwaan et al., 2018); however, the extent to which this occurs in practice is unclear. In this study, we investigated how the research community responded to four strongly contradictory replications in the field of psychology by examining post-replication citation patterns for original studies. We observed some ‘progressive’ response patterns in the form of modest active correction (a small decline in favourable citations and a small increase in unfavourable citations), but also even more prominent ‘regressive’ response patterns in the form of unchallenged belief perpetuation (sustained levels of favourable citations and few unfavourable citations, particularly in the Baumeister case) and considerable citation bias (neglecting to cite the replication study; in all cases aside from Sripada). When authors cited the original study favourably despite the replication result, only half of the articles provided any explicit counter arguments in defence of the original study (an explicit defence). Overall, these findings are consistent with prior observations that favourable citation patterns appear relatively unperturbed by subsequent publication of contradictory results in studies with lower risk of bias (Tatsioni et al., 2007) or even by full retraction (Budd et al., 1998; Fernandez & Vadillo, 2020), and that positive (supportive) evidence is preferentially cited relative to negative (non-supportive) evidence once a theory gets entrenched, despite overwhelming evidence against it (Bastiaansen et al., 2015; Greenberg, 2009).

It is reasonable to question whether the replication results in these case studies should (rationally) have instigated belief change in the research community and thus triggered a more sizable decline in favourable citations than we observed. It is important to note that the case studies we selected were deliberately chosen because the replication results were superior to the original results in terms of both credibility and evidential value. The replications were pre-registered, multi-laboratory studies with large sample sizes. By contrast, the original studies had much smaller sample sizes, and arguably had a much higher risk of bias, as they were not pre-registered, were performed by single teams, and arose in domains affected by publication bias and other questionable research practices (E. C. Carter et al., 2015; Coles et al., 2019; Vadillo, 2019; Vadillo et al., 2016). Thus, it seems reasonable to suggest that the compelling replication results should have reduced belief in the credibility of the original results. However, one could also contend that if there were some flaw in the replication study that undermined its validity (Brandt et al., 2014; Vazire et al., 2021; Fabrigar et al., 2020), then this may provide justification to continue favourably citing the original study. Indeed, in these particular case studies, the validity of the replication results has been challenged by proponents of the original findings (Table 5; Baumeister & Vohs, 2016; Strack, 2016). As this debate remains far from settled (Coles et al., 2019; Vadillo, 2019), ideally any favourable citation of the original studies should at a minimum be accompanied by co-citation of the replication results and some discussion of the discrepant findings.

@@

We have only been able to gauge reactions to replications to the extent that they are reflected in relatively short-term citation patterns. A potential explanation of the apparently cursory treatment of the replication studies could be that researchers only became aware of them after their own research projects had begun and or even had been completed. If one of the original studies had been a key motivator for one’s own study, then it may be difficult to accommodate the strongly contradictory replication results. It may even be tempting to ignore them, or give them only superficial treatment. Moreover, researchers who are convinced that the replication study has squarely refuted the original, may no longer be interested in doing research on a topic where they see no future potential. Additionally, examination of citation patterns would not detect if there had been a correction effect amongst individuals who would not typically cite the original study, such as students, members of the public, or researchers working in other fields. A contested study may continue to be cited favourably by its proponents who remain working in the field. This will suffice to create belief perpetuation in the published literature, even though other scientists may simply no longer be interested in getting involved with such a strongly contested research topic. Relatedly, we did not examine citation patterns beyond 3-5 years post-replication. Some perspectives envision scientific self-correction unfolding over a much longer time scale (Lauden, 1981; Peterson & Panofsky, 2020) and suggest the process is characterised less by the impact of individual study results and more by the gradual accumulation of converging evidence, gradual revision of theoretical understanding, and/or informal sociological processes (e.g., researchers choosing alternative topics to study). Thus, whilst the current findings may contradict the expectations of a more direct and expedited view of the corrective impact of replication studies (Ioannidis, 2012; Vazire & Holcombe, 2020), they are not necessarily inconsistent with a slower and more indirect process of self-correction. Future research could employ longitudinal designs or older historical case studies in order to evaluate citation patterns unfolding over a longer time scale. Finally, as we only examined citation patterns, this study could not capture other potential responses to replication results, such as changes in research practices.

@@

```{r child = "appendix.Rmd"}
```
